Airline Customer Satisfaction - Group 17

### 2025-04-04 index.Rmd
Updated Chapter 3

---
title: "Airline Customer Satisfaction - Group 17"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float: true
  pdf_document: default
date: "2025-04-02"
---

@@

**Correlation Matrix of Numerical Variables**

@@

We'll try to grouping the features into category of services that could be improved by the airline :

Flight General services : Food and Drink, In-flight Service
Flight Experience : In Flight Wifi Service, In Flight Entertainment
Digital services : Ease of Online Booking, Online Boarding
Comfort services : Seat Comfort, Leg Room Service, Cleanliness

```{r}
data$Flight.Service = rowSums(data[,c('Food.and.Drink','In.flight.Service')],na.rm = TRUE)
data$Flight.Entertainment = rowSums(data[,c('In.flight.Wifi.Service','In.flight.Entertainment')],na.rm = TRUE)
data$Digital.Service = rowSums(data[,c('Ease.of.Online.Booking','Online.Boarding')],na.rm = TRUE)
data$Comfort.Service = rowSums(data[,c('Seat.Comfort','Leg.Room.Service','Cleanliness')],na.rm = TRUE)

plot_histogram(data[ , c("Flight.Service", "Flight.Entertainment", "Digital.Service", "Comfort.Service")], ncol=2,ggtheme = theme_minimal())
```

## Data Tranformation <a name="2.4"></a>

```{r}
plot_histogram(data, ncol=2,ggtheme = theme_minimal())
```

We can see that the data is not normally distributed, therefore we need to normalize the data.

@@

Our target is the satisfaction rate of the airlines services, therefore we need to identified how each variable are correlated to the satisfaction

Satisfaction for new features :

```{r}
data_subset <- data %>%
  select(Satisfaction, Flight.Service, Flight.Entertainment, Digital.Service, Comfort.Service)

@@

@@

@@

>>>>

---
title: "Airline Customer Satisfaction - Group 17"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float: true
  pdf_document: default
date: "2025-03-27"
---

@@

@@

Next, We group the features into category of services that could be improved by the airline:

**Flight General Service:** On-board Service, Food and Drink, In-flight Service

**Flight Entertainment:** In Flight Wifi Service, In Flight Entertainment

**Pre-flight Service:** Ease of Online Booking, Check-in Service, Online Boarding

**Comfortability:** Seat Comfort, Leg Room Service, Cleanliness

```{r}
data$In.flight.service = rowSums(data[,c('On.board.Service','Food.and.Drink','In.flight.Service')],na.rm = TRUE)
data$Flight.Entertainment = rowSums(data[,c('In.flight.Wifi.Service','In.flight.Entertainment')],na.rm = TRUE)
data$Pre.flight.service = rowSums(data[,c('Ease.of.Online.Booking','Check.in.Service','Online.Boarding')],na.rm = TRUE)
data$Comfortability = rowSums(data[,c('Seat.Comfort','Leg.Room.Service','Cleanliness')],na.rm = TRUE)

plot_histogram(data[ , c("In.flight.service", "Flight.Entertainment", "Pre.flight.service", "Comfortability")], ncol=2,ggtheme = theme_minimal())
```

Combining several category of services into one variable is useful to see the overall satisfaction of the airline services in numerous area of improvement. For example, as we can see above that in general the passengers found that the flights they are taking are relatively comfortable. However, the pre-flight service and flight entertainment category received mixed reviews and can be considered as an improvement area for the airline, where further analysis could be done to see which of the element of those services are the most dissatisfying for the passengers.

**DARI FEATURE ENGINEERING KE SECTION SELANJUTNYA HARUS CROSSCHECK, BEBERAPA VARIABEL BARU BERUBAH RANGE DAN PENAMAANNYA**

## Data Tranformation <a name="2.4"></a>
Data transformation plays a key role to make sure all data is in the same scale and distribution. we will start  to take a look on our data distribution.

library(ggplot2)
library(DataExplorer)

# Set default colors for histogram bars
update_geom_defaults("bar", list(fill = "steelblue", color = NA))

# Generate histogram with updated colors
plot_histogram(data, 
               ncol = 2, 
               ggtheme = theme_minimal() + theme(strip.background = element_blank(),
                                                 strip.text = element_text(color = "black")))

```
This part is to undertand the distribution of data before it is used. As we can see in the histogram, the data is not normally distributed. Therefore, we need to normalize the data.
There are several concerning variables in the dataset that are part of main focus in the data transformation process.
<ul>
<li>**arrival.delays** and **departure.delays** have highly-right skewed distribution where most values are concentrated to zero.</li>
<li>**flight.distance** has a right-skewed distribution concentrated between 500 and 1500, with a long tail up to 4000.</li>
<li>**age** has a right-skewed distribution concentrated between 20 and 50, with a long tail up to 80.But this variable is less concerning compared to the previous ones.</li>
</ul>
In the following step we will normalize the data by applying log transformation to the concerning variables.

@@

Our target is the satisfaction rate of the airlines services, therefore we need to identify how each variable are correlated to the satisfaction. This is important to understand how each variable's behavior toward main objective of the analysis.
Following is the behavior of each variables towards satisfaction rate.

```{r}
plot_bar(data,by='Satisfaction',ncol=2)
```
According to the graphic above, we can see that:
<ul>
<li>**Gender**: Both male and female customers have a higher proportion of "Neutral or Dissatisfied" responses compared to "Satisfied" ones. This indicates there is a room of improvement to increase satisfied number</li>
<li>**Customer Type**: Returning customers exhibit a higher proportion of "Satisfied" responses than first-time customers, who seem mostly "Neutral or Dissatisfied". This might indicate that familiarity with the company positively influences satisfaction.</li>
<li> **Type of Travel**: Business travelers show a higher proportion of "Satisfied" responses compared to personal travelers.</li>
<li> **Class**: Business class has the highest proportion of "Satisfied" responses, whereas Economy and Economy Plus lean more towards "Neutral or Dissatisfied." This indicates higher satisfaction among premium customers. </li>
<li> **Age Range**: Older and middle-aged customers show higher satisfaction levels compared to youth and children, who lean towards dissatisfaction.</li>
<li> **Distance Group**:Longer trips seem to lead to higher satisfaction levels, which could indicate that services during these trips meet or exceed customer expectations. </li>
<li> **Delay Status**: the behavior of satisfaction between on-time and delayed status is not significantly different </li>
</ul>

```{r}
data_subset <- data %>%
  select(Satisfaction, In.flight.service, Flight.Entertainment, Pre.flight.service, Comfortability)

@@

```{r}
loadings_df <- data.frame(
  variable = rownames(loadings)[indices],
  loading = loadings$PC1[indices]
)

# Plot
ggplot(loadings_df, aes(x = reorder(variable, -loading), y = loading, fill = loading > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Top Loadings for PC1",
       x = "Variable",
       y = "Loading Value") +
  scale_fill_manual(values = c("steelblue", "tomato")) +
  theme_minimal()
```

Analyzing top contributor in PC1-PC12

```{r}
loadings <- as.data.frame(results$rotation)
loadings$Variable <- rownames(loadings)

# Top contributors for PC1–PC12
top_vars <- apply(abs(loadings[, 1:12]), 2, function(x) names(sort(x, decreasing = TRUE)[1:5]))
print(top_vars)
```
PC1 – In-Flight Amenities & Comfort
Driven by Flight.Entertainment, Comfortability, In.flight.Entertainment, In.flight.service, and Cleanliness. This component captures the overall in-flight experience focused on comfort and entertainment.

PC2 – Digital & Booking Experience
Influenced by Ease.of.Online.Booking, Pre.flight.service, and In.flight.Wifi.Service. It represents passengers’ interaction with digital services before and during the flight.

PC3 – Operational Onboard Service
Dominated by In.flight.Service, Baggage.Handling, and Seat.Comfort, this component reflects service reliability and seat-related comfort during the flight.

PC4 – Flight Timing & Boarding
Characterized by Arrival.Delay, Departure.Delay, and Online.Boarding. It relates to punctuality and boarding process efficiency.

PC5 – Passenger Profile & Journey
Shaped by Online.Boarding, Flight.Distance, Age, and Gate.Location, this component reflects demographic and journey-specific traits.

PC6 – Schedule & Gate Convenience
Led by Departure.and.Arrival.Time.Convenience, Gate.Location, and Age, this component captures how convenient flight times and gate accessibility are for passengers.

PC7 – Pre-Flight & Cabin Space
Defined by Check.in.Service, Leg.Room.Service, and Flight.Distance, combining service before boarding with cabin comfort.

PC8 – Age-Based Travel Patterns
Mainly driven by Age, Flight.Distance, and Check.in.Service, this component shows how age may influence travel experience and expectations.

PC9 – Comfort in Long Trips
Includes Leg.Room.Service, Flight.Distance, and Comfortability, highlighting comfort needs for longer or more frequent flights.

PC10 – In-Flight Consumables
Composed of On.board.Service, In.flight.service, and Food.and.Drink, focusing on passengers’ satisfaction with in-flight consumables.

PC11 – Delays & Terminal Logistics
Influenced by Arrival.Delay, Departure.Delay, and Gate.Location, this component represents disruptions tied to flight and gate logistics.

PC12 – Check-In & Time Coordination
Driven by Departure.and.Arrival.Time.Convenience and Check.in.Service, highlighting the efficiency of early-stage travel processes.

```{r}
screeplot(results, type = "lines")
```

**PCA with One Hot Encoding**

```{r}
library(caret)

# Use the standardized data
data_standardized <- data %>%
  select(-c(ID, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
            Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
            Flight.Distance.log.z,Departure.Delay.Duration, Arrival.Delay.Duration,
            Age.log, Age.log.z))

dummies <- dummyVars(~ ., data = data_standardized)
data_encoded <- predict(dummies, newdata = data_standardized)

data_encoded_standardized <- scale(data_encoded, center = TRUE, scale = TRUE)

results <- prcomp(data_encoded_standardized)
summary(results)
```

When PCA was applied to numeric-only variables, the first 12 components explained over 90% of the variance, indicating strong inter-variable correlation and compact data structure.
However, after including one-hot encoded categorical variables, the number of components required to achieve the same level of variance increased to 20. This suggests that categorical variables contribute additional but more dispersed variance, requiring more dimensions to capture the overall structure.

@@

### Model Preparation <a name="3.2.1"></a>

```{r}
# Create copy of dataset for the purpose of classification only, convert to factor
data_clsf <- data %>%
      mutate(across(-c(ID, Age, Flight.Distance, Departure.Delay, Arrival.Delay, 
                   Departure.Delay.Duration,
                   Arrival.Delay.Duration, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
                   Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
                   Flight.Distance.log.z, Age.log, Age.log.z), as.factor))

# Encode Target Variable
data_clsf$Satisfaction <- ifelse(data_clsf$Satisfaction=="Satisfied",1,0)
data_clsf$Satisfaction <- factor(data_clsf$Satisfaction, levels = c(0, 1))

# For experiment, create subset of untransformed data, data with log, and data with normalization
data_clsf_ori <- data_clsf %>%
  select(-c(ID, Arrival.Delay.Duration.log, Arrival.Delay.Duration.log.z,
            Departure.Delay.Duration.log, Departure.Delay.Duration.log.z, Flight.Distance.log,
            Flight.Distance.log.z, Age.log, Age.log.z))

data_clsf_log <- data_clsf %>%
  select(-c(ID, Arrival.Delay.Duration.log.z, Departure.Delay.Duration.log.z, Flight.Distance.log.z,
            Departure.Delay.Duration, Arrival.Delay.Duration, Age.log.z, Arrival.Delay.Duration,
            Departure.Delay.Duration, Flight.Distance, Age))

data_clsf_norm <- data_clsf %>%
  select(-c(ID, Arrival.Delay.Duration.log,
            Departure.Delay.Duration.log, Flight.Distance.log,
            Age.log, Arrival.Delay.Duration,
            Departure.Delay.Duration, Flight.Distance, Age)) 
```

**Start with Untransformed Dataset**

**Split Dataset Training Testing**

```{r}
set.seed(46748717)

library (rsample)

proportion <-  0.7
split <- initial_split(data_clsf_ori, prop = proportion)
training <- training(split)
testing <- testing(split)
```

### Training a decision tree model on the data <a name="3.2.2"></a>

```{r}
model <- C50::C5.0(Satisfaction ~., 
                   data = training)
summary(model)
```

```{r}
print("Confusion matrix based on testing data")
pred.test <- predict(model, testing)
gmodels::CrossTable(testing$Satisfaction, pred.test,
                    prop.chisq = FALSE,
                    prop.c = FALSE,
                    prop.r = FALSE,
                    prop.t = FALSE,
                    dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy

accuracy = (21498 + 15676) / 38964
print(accuracy)
```

Accuracy already high with 95.4%

**Try with dataset that already log transformed**

```{r}
set.seed(46748717)

library (rsample)

proportion <-  0.7
split <- initial_split(data_clsf_log, prop = proportion)
training <- training(split)
testing <- testing(split)

model <- C50::C5.0(Satisfaction ~., 
                   data = training)
summary(model)

print("Confusion matrix based on testing data")
pred.test <- predict(model, testing)
gmodels::CrossTable(testing$Satisfaction, pred.test,
                    prop.chisq = FALSE,
                    prop.c = FALSE,
                    prop.r = FALSE,
                    prop.t = FALSE,
                    dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

The result is almost identical since the algorithm determine the transformed data is not have significant importance. Therefore the improvement will focus on tuning the parameter of the model.

### Model Improvement <a name="3.3.3"></a>

**Boosting**

```{r}
library(C50)
library(gmodels)

set.seed(46748717)

proportion <-  0.7
split <- initial_split(data_clsf_ori, prop = proportion)
training <- training(split)
testing <- testing(split)

model_boost <- C5.0(Satisfaction ~.,
                    data = training,
                    trials = 10)

# Predicting on the testing data
print("Confusion matrix based on testing data (boosting)")
pred.test <- predict(model_boost, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy with Booster

accuracy = (21229 + 15816) / 38964
print(accuracy)
```

Accuracy didnt increase significant

**Assigning Cost to Mistake**

More costly to predict dissatisfied customer satisfied, cost higher in False Positive

```{r}
# Specifying the cost matrix
cost.matrix <- matrix(c(NA, 1,  # FN costs of predicting "Disatisfied" whereas actual value is "Satisfied"
                        2, NA), # FP costs of predicting "Satisfied" whereas actual value is "Disatisfeid"
                      nrow = 2, 
                      ncol = 2,
                      byrow = FALSE)
rownames(cost.matrix) <- colnames(cost.matrix) <- c(1, 0)

# Estimating the model with the cost matrix
model.cost <- C5.0(Satisfaction ~., 
              data = training,
              costs = cost.matrix)

print("Confusion matrix based on testing data (costs")
pred.test <- predict(model.cost, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy with Booster

accuracy = (21755  + 15087 ) / 38964
print(accuracy)
```

Accuracy decrease slightly

```{r}
# Calculating Decrease in False Positive

decrease_fp = (495   - 238  ) / 495
print(decrease_fp)
```

False positive decrease by 51.9%, from 495 to 238

**Random Forest**

```{r random forest}
data_randfor <- data_clsf_ori %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration, Flight.Distance, Age))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix based on testing data (costs")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy with Random Forest

accuracy = (21410   + 15793  ) / 38964
print(accuracy)
```

Accuracy increase 95.4%

**Using Ranger to implement random forest to the database with high cardinality**

```{r}
# install.packages("ranger")
library(ranger)
```

```{r}
proportion <-  0.7
split <- initial_split(data_clsf_ori, prop = proportion)
training <- training(split)
testing <- testing(split)

model.ranger <- ranger(Satisfaction ~., 
                       data = training,
                       num.trees = 500, # how many trees should be grown?
                       mtry = 2, # how many variables to sample at each split?
                       replace = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix based on testing data (costs")
pred.test <- predict(model.ranger, testing)$predictions
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r}
# Calculating Accuracy with Ranger

accuracy = (21412   + 15739  ) / 38964
print(accuracy)
```

Accuracy didnt increase significant

### Variable Importance <a name="3.4"></a>

```{r}
library(randomForest)

importance <- importance(model.forest)
print(importance)
```

Plot the importance

```{r varimp-plot, fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

High satisfaction on :
- Type of travel : Business
- Class : Business
- Customer Type : Returning

Type of Travel : Business

```{r fig.width=10, fig.height=8}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Type.of.Travel == "Business"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Type.of.Travel))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Type of Travel Business")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))

```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Class : Business

```{r}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Class == "Business"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Class))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Class Business")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Customer Type : Returning

```{r fig.width=10, fig.height=8}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Customer.Type == "Returning"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Customer.Type))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Customer Type Returning")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Low Satisfaction on :
- Type of Travel Personal
- Class Economy and Economy Plus
- Customer Type First Time

Type of travel personal

```{r}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Type.of.Travel == "Personal"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Type.of.Travel))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Type of Travel Personal")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Class Economy and Economy Plus

```{r}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Class %in% c("Economy", "Economy Plus")), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Class))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Type of Class Economy and Economy Plus")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

Customer Type First Time

```{r}
data_randfor <- data_clsf_ori[ which(data_clsf_ori$Customer.Type %in% "First-time"), ]
data_randfor <- data_randfor %>% select(-c(Arrival.Delay.Duration, Departure.Delay.Duration,
                                           Flight.Distance, Age, Customer.Type))

proportion <-  0.7
split <- initial_split(data_randfor, prop = proportion)
training <- training(split)
testing <- testing(split)

model.forest <- randomForest::randomForest(Satisfaction ~., 
                                           data = training,
                                           ntree = 500, # how many trees should be grown?
                                           mtry = 2, # how many variables to sample at each split?
                                           replace = TRUE,
                                           importance = TRUE) # sampling of cases with or without replacement?

print("Confusion matrix for Random Forest Type of Customer Type First Time")
pred.test <- predict(model.forest, testing)
CrossTable(testing$Satisfaction, pred.test,
           prop.chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           prop.t = FALSE,
           dnn = c("Actual Satisfaction", "Predicted Satisfaction"))
```

```{r fig.width=10, fig.height=8}
varImpPlot(model.forest)
```

@@

### Training Logistic Regression Model <a name="3.3.2"></a>

```{r}
# Using all data
model_log <- glm(Satisfaction ~., 
               data = training, family = "binomial")

summary(model_log)
```

### Interpretation of the Model <a name="3.3.3"></a>

**Interpret the coefficient**

```{r}
exp(coef(model_log))
```

**Visualize the coefficient**

```{r fig.width=10, fig.height=20}
coefs <- summary(model_log)$coefficients
coef_df <- data.frame(
  Variable = rownames(coefs),
  Estimate = coefs[, "Estimate"],
  StdError = coefs[, "Std. Error"],
  p_value = coefs[, "Pr(>|z|)"]
)

# Remove intercept for clarity
coef_df <- coef_df[coef_df$Variable != "(Intercept)", ]

# Plot using ggplot2
library(ggplot2)

ggplot(coef_df, aes(x = reorder(Variable, Estimate), y = Estimate)) +
  geom_point(color = "darkblue") +
  geom_errorbar(aes(ymin = Estimate - 1.96 * StdError,
                    ymax = Estimate + 1.96 * StdError), width = 0.2) +
  coord_flip() +
  labs(title = "Logistic Regression Coefficients",
       x = "Predictor",
       y = "Coefficient Estimate (log-odds)") +
  theme_minimal()
```

**Interpreting Z Value**

```{r}
model_summary <- summary(model_log)

# Extract z-values and convert to data frame
z_df <- data.frame(
  Variable = rownames(model_summary$coefficients),
  Z_value = model_summary$coefficients[, "z value"]
)

# Remove intercept
z_df <- subset(z_df, Variable != "(Intercept)")

z_df
```

```{r  fig.width=10, fig.height=20}
# Plot using ggplot2
library(ggplot2)

ggplot(z_df, aes(x = reorder(Variable, abs(Z_value)), y = Z_value)) +
  geom_col(fill = "steelblue") +
  geom_hline(yintercept = c(-1.96, 1.96), linetype = "dashed", color = "red") + # 95% significance threshold
  coord_flip() +
  labs(title = "Z-values from Logistic Regression",
       x = "Predictor Variable",
       y = "Z-value (Wald Statistic)") +
  theme_minimal()
```

### Model Evaluation <a name="3.3.4"></a>

```{r}
pred.test <- predict(model_log, newdata = testing, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
testing$Satisfaction <- factor(testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

confusionMatrix(pred.test.result, testing$Satisfaction)
```

Accuracy 93.7% with sensitivity 95.2% and specificy 91.9%

**Interpret ROC and AUC**

```{r}
library(pROC)

# ROC curve and AUC
roc_obj <- roc(testing$Satisfaction, pred.test)
plot(roc_obj, col = "blue", main = "ROC Curve")
auc(roc_obj)
```

```{r}
auc(roc_obj)
```

```{r}
plot(model_log)
```

### Improving Model Performance

**Using significant variables only**

p-value < 0.05

```{r}
training_signif <- training %>% subset(select = -c(Departure.Delay, Arrival.Delay, Flight.Distance, 
                                                  In.flight.Wifi.Service, Food.and.Drink, 
                                                  In.flight.Entertainment, Age_Range, Distance_Group,
                                                  In.flight.service, Flight.Entertainment, 
                                                  Comfortability, Pre.flight.service, 
                                                  Flight.Distance.log.z, Age.log.z))

model_log_signif <- glm(Satisfaction ~., 
               data = training_signif, family = "binomial")

summary(model_log_signif)

```

AIC increase to 44749

```{r}
pred.test <- predict(model_log_signif, newdata = testing, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
testing$Satisfaction <- factor(testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

confusionMatrix(pred.test.result, testing$Satisfaction)
```

The accuracy decrease

ANOVA test :

```{r}
anova(model_log, model_log_signif, test="Chisq")
```

Model full siginificantly better than model with significant only

**Adjusting Threshold**

```{r}
coords(roc_obj, "best", ret = "threshold")
```

```{r}
pred.test <- predict(model_log, newdata = testing, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5012364, "Satisfied", "Neutral or Dissatisfied")
testing$Satisfaction <- factor(testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

confusionMatrix(pred.test.result, testing$Satisfaction)
```

Performance not increase significantly

**Dimensionality Reduction**

```{r}
training_pca <- training %>%
  mutate(across(everything(), ~ if (all(!is.na(as.numeric(as.character(.x))))) 
    as.numeric(as.character(.x)) else .x)) %>%
  select(where(is.numeric)) %>%
  scale(center = TRUE, scale = TRUE)

pca_results <- prcomp(training_pca)
summary(pca_results)
```

12 var already achieve 91%

```{r}
pca_data <- as.data.frame(pca_results$x[, 1:12])
pca_data$Satisfaction <- training$Satisfaction

model_log_pca <- glm(Satisfaction ~., 
               data = pca_data, family = "binomial")

summary(model_log_pca)
```

AIC higher 80736

```{r}
# 1. Convert test data to numeric and scale
test_pca_scaled <- testing %>%
  mutate(across(everything(), ~ if (all(!is.na(as.numeric(as.character(.x))))) 
    as.numeric(as.character(.x)) else .x)) %>%
  select(where(is.numeric)) %>%
  scale(center = attr(training_pca, "scaled:center"), 
        scale = attr(training_pca, "scaled:scale"))  # Use training scale

# 2. Apply PCA transformation
test_pca <- as.data.frame(predict(pca_results, newdata = test_pca_scaled)[, 1:12])

# 3. Add target variable
test_pca$Satisfaction <- testing$Satisfaction

# 4. Predict using the model trained on PCA components
pred.test <- predict(model_log_pca, newdata = test_pca, type = "response")

# Convert to class labels
pred.test.result <- ifelse(pred.test > 0.5, "Satisfied", "Neutral or Dissatisfied")
testing$Satisfaction <- factor(testing$Satisfaction, levels = c("Satisfied", "Neutral or Dissatisfied"))
pred.test.result <- factor(pred.test.result, levels = c("Satisfied", "Neutral or Dissatisfied"))

confusionMatrix(pred.test.result, testing$Satisfaction)
```

Accuracy decrease using PCA

# Interpretation <a name="4"></a>

## Result Visualization <a name="4.1"></a>

## Recommendation <a name="4.2"></a>

## Further Improvement for Better Recommendation <a name="4.3"></a>

## Conclusion & Limitations <a name="4.4"></a>

## References <a name=""></a>
